---
title: "Carbon Intensity API"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

### Imports

```{r}
library(httr)
library(jsonlite)
library(tidyverse)
library(lubridate)
```

### Current Carbon Intensity 

The first request will get the carbon intensity for current half-hour period:  

```{r}
# Make the request
r <- GET('https://api.carbonintensity.org.uk/intensity')
```

The response body is in JSON format, which we can visualise with `prettify()`

```{r}
prettify(content(r, 'text'))
```

To parse the results and convert it to a data frame, we first need to convert the JSON data to an R object. The `fromJSON()` function can be used to convert from JSON to a number of objects automatically. In this case, it will convert the response to a list.

```{r}
# Parse the results from JSON 
parsed <- fromJSON(content(r, 'text'))

class(parsed)
```

There is only one element in this list, `data`, which is a *nested* data frame: the `intensity` column is itself a data frame with three columns: `forecast`, `actual` and `index`. 

```{r}
print(names(parsed$data))
print(names(parsed$data$intensity))
```

We can use `unnest()` to flatten this to a single, unnest data frame wiht 5 columns.

```{r}
# Unnest the intensity data 
df <- parsed$data %>% unnest(cols = intensity)
df
```

Now we will wrap this in a function: 

```{r}
get_carbon_intensity <- function(){
  r <- GET('https://api.carbonintensity.org.uk/intensity')
  parsed <- fromJSON(content(r, 'text'))
  df <- parsed$data %>% unnest(cols = intensity)
  return(df)
}

get_carbon_intensity()
```

### Carbon Intensity for Past Week

We will write three functions:

- Retrieve and parse CO2 intensity data between two dates
- Calculate average CO2 intensity for last week
- Identify period of highest CO2 intensity in the last week

The following function retrieves the carbon intensity data between 2 dates. 

```{r}
co2_intensity_data <- function(date1, date2){
  
  # Create URL
  url <- sprintf('https://api.carbonintensity.org.uk/intensity/%s/%s', date1, date2)

  # Make request, parse, retrieve data
  r <- GET(url)
  parsed <- fromJSON(content(r, 'text'))
  data <- parsed$data %>% unnest(intensity)
  
  return(data)
}
```

Next we calculate the average CO2 intensity for the last week. The following function uses the code above to calculate the average CO2 intensity for the past week. It uses `days()` from lubridate, which can be used to add or subtract periods to/from datetimes.

```{r}
avg_co2_intensity_past_week <- function(){
  # Date 7 days ago
  d1 <- today() - days(7)
  d2 <- today()
  
  # Request and parse data
  data <- co2_intensity_data(d1, d2)
  
  return(mean(data$actual))
}

avg_co2_intensity_past_week()
```

The next function returns both the period and the value of the maximum carbon intensity in the past week. 

```{r}
max_co2_intensity_past_week <- function(){
    # Date 7 days ago
  d1 <- today() - days(7)
  d2 <- today()
  
  # Request and parse data
  data <- co2_intensity_data(d1, d2)
  
  # The maximum intensity
  max_intensity <- max(data$actual)
  
  # Filter the data and extract the period of maximum intensity
  # (we use the beginning of the period (`from`))
  max_period <- data %>% 
    filter(actual == max_intensity) %>% 
    pull(from)
    
  return(c(max_intensity, max_period))
}

result <- max_co2_intensity_past_week()
result
```

Some string formatting can be used to better present this result: 

```{r}
print(sprintf("The maximum carbon intensity during the past week was %s gCO2/kWh at %s", result[1], result[2]))
```

### Next 24h Carbon Intensity

The following code retrieves the forecast carbon intensity for the next 24 hours:

```{r}
# Current date time
dt_now <- now() %>% format('%Y-%m-%dT%H:%MZ') # ISO8601 format

# Use sprintf to format the URL 
url <- sprintf('https://api.carbonintensity.org.uk/intensity/%s/fw24h', dt_now)

# Make the request
r <- GET(url)

# Parse results
parsed <- fromJSON(content(r, 'text'))
df <- parsed$data %>% unnest(cols = intensity)

head(df)
```

### Regional Carbon Intensity

The code below retrieves current regional carbon intensities and prints the regions which have the lowest and highest intensities. 

```{r}
# Make the requeast
r <- GET('https://api.carbonintensity.org.uk/regional')

# Parse the JSON response
parsed <- fromJSON(content(r, 'text'))

# parsed$data$regions is a list of length 1
# The following line retrieves this element, which is a nested data frame
regions <- parsed$data$regions[[1]]

# Unnest the intensity data
regions_intensity <- regions %>% unnest(cols = intensity)

# Retrieve the regions with lowest and highest CO2 intensity
min_region <- regions_intensity$shortname[which.min(regions_intensity$forecast)]
max_region <- regions_intensity$shortname[which.max(regions_intensity$forecast)]

# Print results
print(paste("Current time:", now()))
print(paste("Lowest intensity region:", min_region))
print(paste("Highest intensity region:", max_region))
```
