---
title: "Met Office API"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

### Imports

```{r}
library(httr)
library(jsonlite)
library(tidyverse)
library(lubridate)
```

### Retrieving the site list 

In the code below we retrieve the site list and convert the response into a data frame. 

Note that my `API_KEY` is hidden - it is available as an environment variable in a special file called `.Renviron`. You can read more about setting up your R environment variables [here](https://daattali.gitbooks.io/stat545-ubc-github-io/content/bit003_api-key-env-var.html).

```{r}
# Your API key goes here! I am hiding mine :) 
API_KEY = Sys.getenv('MET_API_KEY')

# Getting site list
url <- sprintf('http://datapoint.metoffice.gov.uk/public/data/val/wxfcs/all/json/sitelist?key=%s', API_KEY)

r <- GET(url)
sitelist <- fromJSON(content(r, 'text')) 
sitelist <- sitelist$Locations$Location %>% tibble()
```

We'll now retrieve the Land's End location ID. 

```{r}
# Land's End
sitelist %>% filter(name == "Land's End")
```

The ID for Land's End is 355205. 

### Retrieving the Forecast

```{r}
location_id <- '352205'

url <- sprintf('http://datapoint.metoffice.gov.uk/public/data/val/wxfcs/all/json/%s?res=3hourly&key=%s', location_id, API_KEY)

r <- GET(url)

data <- fromJSON(content(r, 'text'))
```

The `Wx` component of the response describes each of the parameters:

```{r}
data$SiteRep$Wx$Param
```

The forecast is found in the `DV` component. `data$SiteRep$DV$Location$Period$Rep` is a list of 5 data frames, each giving the forecast for the 5 days. Each data frame has 8 rows: the number of 3hr intervals in a day. 

The following code runs a for loop over the data frames to retrieve the forecast. 

```{r}
# Define 
forecast <- numeric()
for (df in data$SiteRep$DV$Location$Period$Rep){
  forecast <- c(forecast, as.integer(df$S))
}

qplot(1:length(forecast), forecast, geom = 'line') + 
  labs(x = "Horizon (3hr intervals)",
       y = "Wind Speed (mph)")
```

### Retrieving the Feels Like Temperature (F) for tomorrow

An additional query parameter can be added to the URL specifiying a time if we just want the forecast for one period. The code below uses this request to get the Feels Like Temperature (F) for 24 hours ahead. 

```{r}
# Specify location ID 
location_id <- '352205'

# Specify time to get forecast for and convert to ISO 8601 format (lubridate)
time <- (now() + hours(24)) %>% format_ISO8601() 

# Putting together the URL
url <- sprintf('http://datapoint.metoffice.gov.uk/public/data/val/wxfcs/all/json/%s?res=3hourly&time=%s&key=%s', location_id, time, API_KEY)

# Making the request and parsing
r <- GET(url)
data <- fromJSON(content(r, 'text'))

# Retrieve the Feels Like Temperature
feels_like_temp <- data$SiteRep$DV$Location$Period$Rep$F
print(sprintf("The feels like temperature this time tomorrow will be: %sC", feels_like_temp))
```
