---
title: "Supervised Learning Introduction"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

In this notebook we will work through a simple example of a supervised learning task. 

We will use the `mpg` dataset.

### Imports

```{r}
library(tidyverse)
library(broom)
```

### Preparing the data

```{r}
df <- mpg %>% 
  select(hwy, displ, cyl, drv, class) %>%
  mutate_if(is.character, as.factor) # Convert the character variables to factors
```

### Splitting into train and test sets

```{r}
set.seed(123)

train <- sample_frac(df, 0.75) # Take 75% of the rows from df
test <- anti_join(df, train) # All rows in df that are NOT in train
```

### Some exploratory analysis


##### Categorical variables with boxplot

```{r}
df %>% 
  pivot_longer(drv:class) %>%
  ggplot() + 
  geom_boxplot(aes(x = value, y = hwy)) + 
  facet_wrap(~ name, scales = 'free') + 
  coord_flip()
```

##### Continuous variables with hexplot

```{r}
df %>% 
  pivot_longer(displ:cyl) %>%
  ggplot() + 
  geom_hex(aes(x = value, y = hwy)) + 
  facet_wrap(~ name, scales = 'free')
```

### Linear model fitting

```{r}
# Fit a linear model: predict hwy with all other variables
linear_model <- lm(hwy ~ ., data = train)

# This function prints the model as a tidy data frame
tidy(linear_model)
```

### Decision tree fitting

```{r} 
# Fit a decision tree: predict hwy with all other variables
dt <- rpart(hwy ~ ., data = train, 
            minsplit = 10,
            minbucket = 2,
            cp = 0.01)
rpart.plot(dt)
```

### Predictions on test set

```{r}
preds_lm <- predict(linear_model, test)
preds_dt <- predict(dt, test)

preds_df <- tibble(actual = test$hwy,
                   linear_model = preds_lm,
                   decision_tree = preds_dt) %>%
  pivot_longer(c(linear_model, decision_tree),
               names_to = 'model',
               values_to = 'predicted')
```

### Error metrics: MSE

```{r}
mse <- function(actual, predicted){
  answer <- mean((predicted - actual)**2)
  return(answer)
}
```

### Error metrics: RMSE

```{r}
rmse <- function(actual, predicted){
  answer <- sqrt(mse(actual, predicted))
  return(answer)
}
```

### Error metrics: MAE

```{r}
mae <- function(actual, predicted){
  answer <- mean(abs(predicted - actual))
  return(answer)
}
```

### Accuracy metrics

```{r}
preds_df %>% 
  group_by(model) %>% 
  summarise(mse = mse(actual, predicted),
            rmse = rmse(actual, predicted),
            mae = mae(actual, predicted))
```

### Plot: predicted vs. actual 

```{r}
preds_df %>% 
  ggplot() + 
  geom_point(aes(x = actual, y = predicted)) + 
  facet_wrap(~model) + 
  geom_abline(intercept = 0, slope = 1, color = 'gray50')
```

### Plot: residuals plot

```{r}
preds_df %>% 
  mutate(residual = predicted - actual) %>% 
  ggplot(aes(x = actual, y = residual)) + 
  geom_point() +
  geom_smooth(method = 'lm') +
  facet_wrap(~model)
```



