---
title: "Pivoting and Facets"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

In this notebook we will investigate **pivoting** and **tidy data**. 

### Imports 

```{r}
library(tidyverse)
```


One of the key tenets of the tidyverse is its reliance on tidy data. Tidy data has the following rules:

- Each variable must have its own column.
- Each observation must have its own row.
- Each value must have its own cell.

Take a look at the following dataset of energy consumption by sector and fuel: 

```{r}
energy <- read_csv('data/energy_consumption_by_sector_fuel.csv')
```

Which rule(s) of tidy data does the dataset violate? 

### Task: use pivot_longer() to make the energy data tidy

```{r}
longer <- energy %>% 
  pivot_longer(Solid_Fuel: Electricity,
               names_to = 'fuel',
               values_to = 'consumption') 
```

### Task: create the plot on the right with ggplot2 and geom_line()

```{r}
longer %>% 
  ggplot() + 
  geom_line(aes(x = Year, 
                y = consumption, 
                color = fuel, 
                linetype = Sector))
```

### Task: use pivot_wider() to return the pivoted energy data frame back to its original

```{r}
longer %>% 
  pivot_wider(names_from = fuel, values_from = consumption)
```

### Task: use pivot_wider() to create a new data frame with the following columns: Sector, fuel, 1970, 1971,…2017

```{r}
longer %>% 
  pivot_wider(names_from = Year, values_from = consumption)
```

### Task: use facet_wrap() to create a bar plot for each sector (see right)

```{r}
# Creating the base plot
plot <- longer %>% 
  ggplot() + 
  geom_col(aes(x = Year, 
               y = consumption,
               fill = fuel)) 

plot + facet_wrap(~Sector)
```

### Task: use the `scales` argument to allow each panel to be freely scaled by consumption

```{r}
plot + facet_wrap(~Sector, scales = 'free_y')
```

### Task: create a bar plot of consumption vs. year, coloured by Sector, faceted by fuel

```{r}
longer %>% 
  ggplot() + 
  geom_col(aes(x = Year, 
               y = consumption,
               fill = Sector)) +
  facet_wrap(~fuel, scales = 'free_y')
```

### Task: use facet_grid() to create a line plot for each fuel and sector (see right)

```{r}
longer %>% 
  ggplot() + 
  geom_line(aes(x = Year, 
               y = consumption)) +
  facet_grid(Sector ~ fuel)
```

### Task: replace facet_grid() with facet_wrap() and set the y scale to be free

```{r}
longer %>% 
  ggplot() + 
  geom_line(aes(x = Year, 
                y = consumption)) +
  facet_wrap(Sector ~ fuel, scales = 'free')
```



