---
title: "Programming"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

### Imports 

```{r}
library(tidyverse)
```


##### Task: write a function addition(x, y) which returns the sum of two arguments

```{r}
addition <- function(x, y){
  return(x + y)
}
addition(1,2)
```

##### Task: Write a function to calculate the interquartile range of a vector

```{r}
iqr <- function(v){
  upper <- as.numeric(quantile(v, 0.75))
  lower <- as.numeric(quantile(v, 0.25))
  return(upper - lower)
}

vector <- rnorm(10000)
iqr(vector)
```

### mpg tasks

##### Task: Calculate the average hwy MPG for a specified manufacturer 

```{r}
average_hwy_mpg <- function(m){
  df <- mpg %>%
    filter(manufacturer == m)
  ans <- mean(df$hwy)
  return(ans)
}

average_hwy_mpg('audi')
```

##### Task: Return the highest cty MPG for a specified class and drv

```{r}
max_cty_mpg <- function(c, d){
  df <- mpg %>%
    filter(class == c,
           drv == d)
  ans <- max(df$cty)
  return(ans)
}

max_cty_mpg('suv', 'f')

```

##### Task: Calculate the correlation between cty MPG and displ for a specified class

```{r}
cor_mpg_displ <- function(c){
  df <- mpg %>%
    filter(class == c)
  
  ans <- cor(df$cty, df$displ)
  
  return(ans)
}

cor_mpg_displ('suv')
```

##### Task: write a function that prints “Even” or “Odd” depending if the input an even/odd number

```{r}
even_or_odd <- function(x){
  
  if (x %% 1 != 0){
    print("Not an integer")
  } else if (x %% 2 == 0){
    print("Even") 
  } else {
    print("Odd")
  }
}

even_or_odd(1521.5)

```

##### Task: Write a function that checks if a letter is a vowel or consonant

```{r}
is_vowel <- function(letter){
  if (letter %in% c('a', 'e', 'i', 'o', 'u')){
    return(TRUE)
  } else{
    return(FALSE)
  }
}

is_vowel('b')

```

##### Task: Using a nested if-else statement, write a function that prints: 

- ”Big and even!” if input is an even number that is greater than 10
- “Big and odd!” if input is an odd number that is greater than 10
- ”Small :(” otherwise

```{r}

number_func <- function(x)
  if (x > 10){
    if (x %% 2 == 0){
      print("Big and even!")
    } else {
      print("Big and odd!") }
  } else {
    print("Small :(")
  }
number_func(11259125)
```

##### Task: Write a function that finds the roots of a quadratic equation given coefficients a, b, c

```{r}
find_roots <- function(a, b, c){
  z = b**2 - 4*a*c
  if (z < 0){
    print("No real solution")
    return()
  }
  
  x1 = (-b + sqrt(z)) / 2*a
  x2 = (-b - sqrt(z)) / 2*a

  return(c(x1, x2))
}

find_roots(1, 10, -5)
```

##### Task: Write a function to change a word to lower case and remove vowels

```{r}
convert_string <- function(word){
  word <- word %>%
    tolower() %>%
    str_remove_all('[aeiou]')
  return(word)
}

convert_string('ESDA')
```

##### add a `engine_category` column to mpg that is “large” for cars with at least a 2L engine displacement (`displ`), and small for all other cars

```{r}
mpg %>% 
  mutate(engine_category = if_else(displ >= 2, 'large', 'small'))
```


##### Task: write a for loop that prints the numbers 1 to 100 (multiples of 7)

```{r}
for (i in 1:100){
  if (i %% 7 == 0){
    print(i)
  }
}
```

### BEIS Headcount Tasks

1. Write a for loop that reads each of the CSV files in the beis_headcount directory

```{r}
data_dir <- 'data/beis_headcount'
files <- list.files(data_dir)
for (f in files){
  df <- read_csv(file.path(data_dir, f))
}
```

2. For each year:

- Total headcount

```{r}
data_dir <- 'data/beis_headcount'
files <- list.files(data_dir)
for (f in files){
  df <- read_csv(file.path(data_dir, f)) %>%
    drop_na(`Total Headcount`)
  hc <- sum(df$`Total Headcount`)
  year <- str_extract(f, "[0-9]+")
  print(year)
  print(hc)
}
```

- Headcount of the Committee for Climate Change
- Headcount of the UK Space Agency 

```{r}
data_dir <- 'data/beis_headcount'
files <- list.files(data_dir)
for (f in files){
  df <- read_csv(file.path(data_dir, f))
  hc <- df %>% 
    filter(`Organisation name` == 'UK Space Agency') %>% 
    pull(`Total Headcount`)
  print(f)
  print(hc)
}
```

- Department with the largest headcount

```{r}
data_dir <- 'data/beis_headcount'
files <- list.files(data_dir)
for (f in files){
  df <- read_csv(file.path(data_dir, f))
  largest <- df %>% 
    filter(`Total Headcount` == max(`Total Headcount`)) %>% 
    pull(`Organisation name`)
  print(f)
  print(largest)
}
```

3. Write a for loop to read all of the CSV files and combine them into a single data frame

```{r}
data_dir <- 'data/beis_headcount'
files <- list.files(data_dir)
all_df <- tibble()
for (f in files){
  df <- read_csv(file.path(data_dir, f))
  all_df <- all_df %>% bind_rows(df)
}
```